Classification of Body Fat based on Measurements and Body Composition

### 2024-12-14 paper/paper.qmd
model section without references

---
title: "Classification of Body Fat based on Physical Measurements and Body Composition"
subtitle: "A comparitive analysis of methods of classifying body fat"
author: 
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/aamishi/ImprovedBodyFatClassification"
date: today
date-format: long
abstract: "BMI values fail to acknowledge the different body composition across both sexes. Using the Waist-Hip-Ratio is more accurate at estimating body fat levels using simple methods at home. This paper compares how this model performs against actual body composition such as isolated fat mass. Using this method makes self-identification of obsesity more accessible and preventive measures can be taken earlier."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Data Section [#paper-data-section]

@@

The primary motivator of this paper was the diverse body compositions that were classified under the same BMI category. The graph [@tbl-bmi-cats-bf] shows the diverse values of Body Fat Percentages classified in the same BMI category.
```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-bmi-cats-bf
#| tbl-cap: "BMI vs. Fat Mass by BMI Category"

@@

# Data {#sec-data}

## Overview
#### General Statement
The data used in this paper comes from two distinct sources, each serving a specific purpose:

2. **Body Composition Dataset**  
This dataset, referred to as the Body Composition dataset for the purposes of this paper, was collected for estimating a new body fat measurement technique by Isaac Kuzmar et al. [@Kuzmar2020]. It is used to calculate fat percentages based on actual body composition and serves as a comparison for the model’s predictions.

1. **BodyM Dataset**  
   This dataset, obtained from Amazon Web Services (AWS) [@AWS_Bodym], is used as training data for the model to predict body fat categories. It includes silhouettes of real test subjects to capture accurate body compositions. These silhouettes were collected to support the estimation of bodily measurements using Machine Learning techniques. However, the silhouettes are not used in this paper.

The BodyM Dataset, referred to as the Body Measurements dataset throughout this paper, was collected by Ruiz et al. [@Ruiz2022] with a focus on underrepresented body types in the estimation of fat and its subsequent health risks. The primary data captured in this collection consists of the front and lateral silhouettes of approximately [2000] test subjects. These silhouettes were then converted into black-and-white images for use in their augmentation model. The dataset includes X male and Y female subjects, aged A to B. The body measurements in this dataset were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types. The S3 package [@reference] included three datasets: Training, Test A, and Test B. For the purposes of this paper, only the Training dataset is used. The visual images of the test subjects were photographed and 3D-scanned by lab technicians.

@@

For this paper, the BodyM Dataset is used to estimate the category of body fat that a person carries. To effectively differentiate between the various body types, the model takes into account measurements that could indicate actual body compositions. Gender, height, and weight are used as basic predictors for the model. However, according to [@reference, WHO], the waist-to-hip ratio (WHR) is a more accurate predictor of body fat levels. Based on [@reference], the two sexes exhibit different tendencies for excess body fat accumulation, also known as adipose fat. In males, excess fat tends to accumulate in the abdominal region, whereas in females, fat is more commonly stored in the hips. This ratio is indicative of fat distribution, and a higher WHR can be a precursor indicator of obesity [@reference].

To better account for body composition in relation to height, a new variable called height-to-hip ratio (height_hip_ratio) was introduced. As noted by [@reference], two individuals with the same WHR could differ in their fat category depending on height. For instance, a taller person with a higher WHR may face a lower risk of developing cardiovascular issues than a shorter person with the same WHR [@reference].

Finally, to enhance the model’s quality, ankle and wrist circumference measurements were also included. These measurements are particularly important for determining boundary values for fat classification [@reference].

@@

Both models employ `fat_percentage_category` as their independent variable or estimand. The reason for chosing a category over an actual value was the difference in predictors across the two models. Both models use different predictors to estimate `fat_percentage_category` so in order to produce useful comparisons against common predictors such as sex, height and weight, I used a categorical variable.

@@

cleaned_data_model$bmi_category <- factor(cleaned_data_model$bmi_category, 
                            levels = c("Underweight", "Normal", "Overweight", "Obese"))
ggplot(cleaned_data_model, aes(x = gender, y = height_hip_ratio, fill = gender)) +
  geom_boxplot() +
  ylab("Waist-to-Hip Ratio") +
  xlab("Gender") +
  labs(title = "Distribution of Waist-to-Hip Ratio by Gender") +
  theme_minimal() +
  facet_wrap(~bmi_category) + 
  scale_fill_manual(values = c("female" = "lightblue", "male" = "lightgreen"))
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
# filtered_data <- subset(cleaned_data_model, bmi_category %in% c("Overweight", "Obese"))

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(body_mass_model, aes(x = weight, y = height, color = fm)) +
  geom_point(size = 3) +
  scale_color_viridis_c() + # Use a continuous color scale
  labs(title = "Height vs Weight with Fat Mass",
       x = "Weight",
       y = "Height",
       color = "Fat Mass (fm)") +
  theme_minimal()
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(body_mass_model, aes(x = weight, y = height, color = fm, size = fm)) +
  geom_point(alpha = 0.7) +
  scale_color_viridis_c() +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Height vs Weight with Fat Mass",
       x = "Weight",
       y = "Height",
       color = "Fat Mass (fm)",
       size = "Fat Mass (fm)") +
  theme_minimal()
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(body_mass_model, aes(x = bmi, y = fm, color = gender)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ gender) +
  labs(title = "Fat Mass vs BMI by Gender",
       x = "BMI",
       y = "Fat Mass (fm)",
       color = "Gender") +
  theme_minimal()
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
body_mass_long <- body_mass_model %>%
  pivot_longer(cols = c(fm, bmi), names_to = "measure", values_to = "value")

@@

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

@@

# Model

I aimed to provide a categorization of body fat levels in an individual based on simple measurements. By using measurements such as height, weight, waist-to-hip ratio (WHR), and ankle and wrist circumference, I sought to capture a person’s body composition as accurately as possible. I expected a positive relationship between excess body fat levels and WHR. Additionally, height is an important factor to consider in capturing fat distribution—specifically, a shorter person with a high WHR would be at a higher risk of obesity-related diseases.

Men are more likely to accumulate fat in the abdominal region, while women tend to store fat in the hip area. To capture natural body shapes and leanness, I also considered other factors that influence body composition, such as height-hip ratio that would capture the distribution of fat
.
## Model General Overview:

This paper uses 2 models to model the two datasets. The first model, the Body Measurements Model, would be used to predict fat percetage categories using the measurements of person’s body that can be obtained using a measuring tape. The second model, the Body Composition Model, is also designed to access the relation between different bodiliy compositions such as total body weight, excess fat weight, lean muscle mass in kilograms.

I used the packages [] to do []. Both models employ Multilevel Regression [] as the Fat Percentage Category has multiple categories including: essential, athlete, normal, high.

### Model 1:

In this model, the dependent variable is fat_percentage_category. The model is set up as follows

\begin{align}
\text{log}\left(\frac{P(y = j)}{P(y = \text{ref})}\right) = \beta_{0j} + \beta_{1j} \cdot \text{height} + \beta_{2j} \cdot \text{gender} + \beta_{3j} \cdot \text{waist\_hip\_ratio} + \beta_{4j} \cdot \text{height\_hip\_ratio}
\end{align}

<!-- Where: -->
<!-- \begin{align} -->
<!-- - \( y \) is the **fat percentage category** (e.g., categories like 'low', 'medium', 'high'). -->
<!-- - \( P(y = j) \) is the probability that the fat percentage falls into category \( j \). -->
<!-- - \( P(y = \text{ref}) \) is the reference category. -->
<!-- - \( \beta_{0j}, \beta_{1j}, \dots \) are the coefficients to be estimated for each category \( j \). -->
<!-- \end{align} -->

Where:
- \( y \) is the **fat percentage category**. The categories include low, moderate, high, and extreme. These equate low, moderate, high and extreme levels of body fat.
- \( P(y = j) \) is the probability that the fat percentage falls into category \( j \).
- \( P(y = \text{ref}) \) is the reference category.
- \( \beta_{0j}, \beta_{1j}, \dots \) are the coefficients to be estimated for each category \( j \).

@@

It is incredibily challenging to differentiate the body composition of a person based soley on their total weight, height and weight only. However, to increase the [goodness] of the classification, this model uses the waist_to_hip ratio, which is especially an important indicator for fat accumualtion across both genders. The model also uses height_hip_ratio to account for different body types. [this sentence might go in the datasection] As height is an important factor to evaluate the collection of fat in latitdunial(how tf do you spell it) places on a person’s body, this is an important factor to consider. However, using height simply does not effectvely communicate how a person’s body is composed latitudinally. So I chose to employ the height_hip_ratio ratio to provide more classification between body types. Additional features such and ankle and wrist are used to further differentiate between the categories.

@@

Assumptions:
- this model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.

@@

- The dataset does not include the age of the particapants. while a general age range is provided, such as 18-X years old, different age groups could differ in how they are categorised.
- This model would not be able to differentiate between age groups and hence would work poorly if this information is included.
-

### Model validation

---

### Model 2:

The model fits a multinomial logistic regression model to predict the fat percentage category (`fat_percentage_category`) based on the following predictors:

- `height` (a continuous variable representing the individual's height),
- `gender` (a categorical variable representing the individual's gender),
- `fm` (a continuous variable representing fat mass or a related measure).

The model is defined as follows:

\begin{align}
P(Y = k \mid X_1, X_2, X_3) = \frac{e^{\beta_{0k} + \beta_{1k} X_1 + \beta_{2k} X_2 + \beta_{3k} X_3}}{1 + \sum_{m=1}^{K-1} e^{\beta_{0m} + \beta_{1m} X_1 + \beta_{2m} X_2 + \beta_{3m} X_3}}
\end{align}

@@

- \( P(Y = k \mid X_1, X_2, X_3) \) is the probability of observing the \( k \)-th category of `fat_percentage_category` given the values of `height` (\( X_1 \)), `gender` (\( X_2 \)), and `fm` (\( X_3 \)).
- \( \beta_{0k}, \beta_{1k}, \beta_{2k}, \beta_{3k} \) are the estimated coefficients for each category \( k \) of the outcome variable, where the reference category is typically set to 0.
- \( K \) is the total number of categories in the dependent variable `fat_percentage_category`.

This model allows us to assess the relationship between the categorical outcome (`fat_percentage_category`) and the continuous and categorical predictors (`height`, `gender`, `fm`), estimating the probabilities for each category of fat percentage based on the input predictors.

@@

@@

Assumptions:
- this model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.

@@

- The dataset does not include the age of the particapants. while a general age range is provided, such as 18-X years old, different age groups could differ in how they are categorised.
- This model would not be able to differentiate between age groups and hence would work poorly if this information is included.
- This model is also built on data that was generated by the ABS [] simulator so it may be inherently working with an error.

### Model validation:

Model out-of-sample testing:

The dataset provided by BodyM already had a trainining and testing split. I used the Test A to test the accuracy of the model. As the method of collection was similar to the training data, the potential errors in the model could only arise from the fact that it is unseen data.

More elaboare explanation, tables and figures can be found at [].

RSME:
I also computed the RSME to evaluate the model’s performance as a lower RSME indicates accuracy in predicted vs actual fat percentage categories.

# Results
Our results are summarized in @tbl-measurementmodelresults and @tbl-bodymass-modelresults. Our model's results and interpretation:

@@

# Model details {#sec-model-details}

>>>>

---
title: "Classification of Body Fat based on Measurements and Body Composition"
subtitle: "Comparitive analysis of Body Fat classification based on BMI and Fat Percentage"
author: 
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/aamishi/ImprovedBodyFatClassification"
date: today
date-format: long
abstract: "BMI values fail to acknowledge the different body composition across both sexes. Using the Waist-Hip-Ratio is more accurate at estimating body fat levels using simple methods at home. This paper compares how this model performs against actual body composition such as isolated fat mass. Using this method makes self-identification of obsesity more accessible and preventive measures can be taken earlier."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

# Data Section {#paper-data-section}

@@

The primary motivator of this paper is the misclassification of diverse body types into the same BMI categories. The graph [@fig-bmi-cats-bf] is a reference point for the main argument of the paper and {shows} how true body fat percentages differ within the same BMI category. As both data sets have different methods of estimating fitness levels, the aim of this paper is to generate a connection between the two data sets. In both data sets, I have created the variable `fat_percentage_category` that the model predict based on entirely different predictors measured in different scenarios.

[x] The primary motivator of this paper was the diverse body compositions that were classified under the same BMI category. The graph [@fig-bmi-cats-bf] shows the diverse values of Body Fat Percentages classified in the same BMI category.
```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-bmi-cats-bf
#| tbl-cap: "BMI vs. Fat Mass by BMI Category"

@@

## Overview
### General Statement
The data used in this paper comes from two distinct sources, each serving a specific purpose:

## **Body Measurements Dataset**  {refer}
This data set, referred to as the Body Measurements data set for the purposes of this paper, was collected for estimating a new body fat measurement technique by Isaac Kuzmar et al. [@Kuzmar2020]. The original paper ....

It is used to calculate fat percentages based on actual body composition and serves as a comparison for the model’s predictions.

This dataset was

## BodyM Data Set
   This dataset, obtained from Amazon Web Services (AWS) [@AWS_Bodym], is used as training data for the model to predict body fat categories. It includes silhouettes of real test subjects to capture accurate body compositions. These silhouettes were collected to support the estimation of bodily measurements using Machine Learning techniques. However, the silhouettes are not used in this paper.

The BodyM Dataset, referred to as the Body Measurements dataset throughout this paper, was collected by Ruiz et al. [@Ruiz2022] with a focus on underrepresented body types in the estimation of fat and its subsequent health risks. The primary data captured in this collection consists of the front and lateral silhouettes of approximately [2000] test subjects. These silhouettes were then converted into black-and-white images for use in their augmentation model. The data set includes X male and Y female subjects, aged A to B. The body measurements in this data set were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types. The S3 package [@reference] included three datasets: Training, Test A, and Test B. For the purposes of this paper, only the Training data set is used. The visual images of the test subjects were photographed and 3D-scanned by lab technicians.

@@

For this paper, the BodyM Dataset is used to estimate the category of body fat that a person carries using circumference measurements. This means the circumferential length of body parts such as the waist and the hips. To build the model, I chose standard predictors such as gender, height and the weight of subjects as the base predictors. However, these predictors are not effective at differentiating body types. Therefore, to effectively construct the distribution of body fat in human body only through measurements, I created two variables called `waist_hip_ration` and `height_hip_ratio`. These variables are used to model the differences in the lateral and horizontal fat distributions. According to WHO [@reference], the waist-to-hip ratio, WHR, is a more accurate predictor of body fat around the abdominal area. Based on the study done in [@reference], the two sexes exhitbit difference tendenices for excess body fat accumulation, also known as the adipose fat. In males, excess fat tends to accumulate in the abdominal region, whereas in females, fat is more commonly stored in the hips. This ratio is indicative of fat distribution, and a higher WHR can be a precursing indicator of obesity [@reference]. In the study done in [place], it is shown that an increase in height lowers the decreases the probability of being classified into a higher BMI category for both genders.
Finally, to enhance the model’s quality, ankle and wrist circumference measurements were also included. These measurements are particularly important for determining boundary values for fat classification [@reference]. Excess fat accumulation at the wrists and ankles could indicate other health risks as seen in [].

@@

Using the above predictor variables, I contructed the variable `fat_percentage_category` as I recruit a multilevel linear regression model to classify the body fat. Initially, I wanted to predict the actual fat percentage, but this proved difficult as ... I decided to use a categorical variable as the estimand to produced useful comparisons against other categorical variables such as sex, height categories and weight categories. Additionally, `fat_percentage_category` was constructed using BMI as a proxy to gauge fat levels in a person. BMI is still widely used in medical scenarios to provide general diagnosis of health levels so I used that as a base to study the interactions between current systems and new offerings of WHR and height_hip_ratio. The relationship between `fat_percentage_category` and BMI was considered to be relatively straightforward to mimick the nature of the current applications of BMI. For BMI, I chose to consider the categories, `underweight`, `normal`, `overweight` and `obese`. These were calulated using the BMI formula:
[insert BMI formula on a new line].

No difference between the genders was considered. Obesity categories using BMI consists of Obese I, Obese II and Obese III as they cause the increasing obesity relating such as diabetes, [ii] and [iii]. However, due to the lack of data points for Obesity II and III (BMI X - X), I chose to categorize all levels of obesity within the same `obesity` category.

The following table shows the relationship between the new `fat_percentage_category` against existing the `bmi_category`.

@@

cleaned_data_model$bmi_category <- factor(cleaned_data_model$bmi_category, 
                            levels = c("Underweight", "Normal", "Overweight", "Obese"))
ggplot(cleaned_data_model, aes(x = gender, y = height_hip_ratio, fill = gender)) +
  geom_boxplot() +
  ylab("Waist-to-Hip Ratio") +
  xlab("Gender") +
  labs(title = "Distribution of Waist-to-Hip Ratio by Gender") +
  theme_minimal() +
  facet_wrap(~bmi_category) + 
  scale_fill_manual(values = c("female" = "lightblue", "male" = "lightgreen"))
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
# filtered_data <- subset(cleaned_data_model, bmi_category %in% c("Overweight", "Obese"))

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(body_mass_model, aes(x = weight, y = height, color = fm)) +
  geom_point(size = 3) +
  scale_color_viridis_c() + # Use a continuous color scale
  labs(title = "Height vs Weight with Fat Mass",
       x = "Weight",
       y = "Height",
       color = "Fat Mass (fm)") +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(body_mass_model, aes(x = weight, y = height, color = fm, size = fm)) +
  geom_point(alpha = 0.7) +
  scale_color_viridis_c() +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Height vs Weight with Fat Mass",
       x = "Weight",
       y = "Height",
       color = "Fat Mass (fm)",
       size = "Fat Mass (fm)") +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(body_mass_model, aes(x = bmi, y = fm, color = gender)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ gender) +
  labs(title = "Fat Mass vs BMI by Gender",
       x = "BMI",
       y = "Fat Mass (fm)",
       color = "Gender") +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
body_mass_long <- body_mass_model %>%
  pivot_longer(cols = c(fm, bmi), names_to = "measure", values_to = "value")

@@

## Measurement of Data

### Measurements Data Set

### Body Composition Data Set

@@

\newpage
# Models

Both models, the Measurements Model and the Body Composition Model have the same independent variable that is based on different proxies. For a gross estimation, the primary proxy for each model against Fat Percentage Category is included in [@tbl-gross-estimation] below. Reasons behind this decision are discussed in greater detail in each model section. Both models are Multilevel Regression Models to predict `fat_percentage_category` into categories of "Low", "Moderate", "High" and "Extreme" levels. I used the `nnet` (@R-nnet) package in R (@citeR) to create both multilevel regression models. Model diagnostics including error checks is included in Appendix [@sec-model-details] along with more elaborate explanations, tables and figures.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-gross-estimation
#| tbl-cap: "Fat Percentage Category against proxies: BMI Category and Actual Fat Percentage"

# Data for the table
gross_estimation <- data.frame(
  Fat_Percentage_Category = c("Low", "Moderate", "High", "Extreme"),
  BMI_Category = c("Underweight", "Normal", "Overweight", "Obese"),
  Male = c("<8%", "8-20%", "21-25%", ">25%"),
  Female = c("<21%", "21-33%", "34-39%", ">39%")
)

# Create the Kable table
kable(
  gross_estimation,
  col.names = c("Fat Percentage Category", "BMI", "Fat Percentage (Male)", "Fat Percentage (Female)"))
```

## Measurements Model
I aimed to provide a categorization of body fat levels in an individual based on simple measurements. By using measurements such as height, weight, waist-hip-ratio (WHR), and ankle and wrist circumferences, I sought to capture a person’s body composition as accurately as possible. I expected a positive relationship between excess body fat levels and WHR. Additionally, height is an important factor to consider in capturing fat distribution — specifically, a shorter person with a high WHR would be at a higher risk of obesity-related diseases. As men are more likely to accumulate fat in the abdominal region, while women tend to store fat in the hip area. To capture natural body shapes and contours, the height-hip-ratio is also a dominant predictor in the model.

In this model, the dependent variable is `fat_percentage_category`. The model is set up as follows

<!-- my model eq -->
\begin{align}
\log \left( \frac{P(Y = k \mid X)}{P(Y = \text{baseline} \mid X)} \right) = \beta_{0k} + \beta_{1k} \cdot \text{height} + \beta_{2k} \cdot \text{gender} + \beta_{3k} \cdot \text{waist\_hip\_ratio} + \beta_{4k} \cdot \text{height\_hip\_ratio}
\end{align}

Where

TODO !!!!!! change this to normal

- $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where \( k \in \{1, 2, \dots, K\} \)) given the predictors $X$.
- $\text{baseline}$ is the reference category (the category of the outcome variable that is omitted in the model, typically the first category).
- $\beta_{0k}$, $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the intercept and the predictors for category \( k \).
- The predictors in this model are `height`, `gender`, `waist_hip_ratio`, and `height_hip_ratio`.

@@

It is incredibly challenging to differentiate the body composition of a person based solely on their total weight and height only. However, to increase the accuracy of the classification, this model uses the `waist_to_hip` ratio, which is especially an important indicator for fat accumulation across both genders. The model also uses `height_hip_ratio` to account for different body types. While height is an important physical attribute, simply using the height variable does not effectively capture how a person’s body is composed horizontally. So, I chose to employ the `height_hip_ratio` ratio to provide more classification between body types. Additional features such and ankle and wrist are used to further differentiate between the categories. As this model uses BMI as a proxy for body fat, using both height and weight as interacting predictors produced no difference in the new predicted categories of `fat_percentage_category` as the same variables were indirectly used as proxies to assign `fat_percentage_category` using the original BMI formula. Therefore, I chose to only consider height. Another reason for opting out of weight as a predictor, was again, the inability of the model to differentiate between fat and muscle mass. So removing weight altogether was successful in the new classification of fat level.

@@

Assumptions:

his model assumes that there is a positive relationship with the higher classification of the fat percentage category and the fat storing tendencies in males and females. Other physical ailments that would cause an increase in the circumference measurements without increasing the overall fat in the body is not considered. This model assumes that females with a higher WHR and males with a higher height-hip ratio are classified higher in the fat categories.

@@

The training data set for this model does not include the age of the participants. This is a huge drawback when it comes to assigning risk labels as aging changes how fat is perceived in the bod. Hence, this model would not be able to differentiate between age groups and would work poorly if this information is included. I also considered estimating actual the fat percentage instead of fat percentage **category**. However, to increase the comparability across both models, I chose to use a categorical variable instead.

## Body Composition Model

The model fits a multinomial logistic regression model to predict the fat percentage category (`fat_percentage_category`). The model is defined as follows under the assumption that "Normal" is the baseline category of fat classification:

\begin{align}
\log \left( \frac{P(Y = k \mid X)}{P(Y = \text{Normal} \mid X)} \right) = \beta_{1k} \cdot \text{age} + \beta_{2k} \cdot \text{height} + \beta_{3k} \cdot \text{gender} + \beta_{4k} \cdot \text{fm}
\end{align}

@@

- $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where (k \in \{\text{Low}, \text{High}, \text{Extreme) given the predictors $X$.
- $\text{Normal}$ is the baseline category (the omitted category), meaning the model compares all other categories to "Normal".
- $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the predictors for category \( k \), relative to the baseline category "Normal".
- The predictors in this model are `age`, `height`, `gender`, and `fm` (which stands for fat mass percentage).

This model allows us to assess the relationship between the categorical outcome (`fat_percentage_category`) and the continuous and categorical predictors (`age`, `height`, `gender`, `fm`), estimating the probabilities for each category of fat percentage based on the input predictors.

@@

The equation for the model expresses the **log-odds** of being in a given category of `fat_percentage_category` with respect to the baseline category "Normal". For each non-baseline category, the model estimates a separate set of coefficients that describe the log-odds of being in that category compared to being in the "Normal" category. The existence of excess fat in a person has been associated with increased obesity related risks. This model primarily focuses on actual fat in a person and is the main predictor variable of interest. I have considered only two other predictors: `gender` and `height` of a person as there is physiological difference in how access fat affects males and females of persons of different height across different ages. So, `age` is also considered as a predictor variable. Additionally, there are different thresholds to classify a person based on the same fat percentage. As seen in the previous model, I want to study the precise difference in estimation of fat levels when weight is not considered holistically. Hence, `gender` and `height` were considered to be comparable with the Body Measurements Models. I had also considered using muscle mass of the subjects. However, the data was skewed toward obese people as this data sets was specifically consisted of people who wanted to improve body image. I decided to omit muscle mass as it interacted with the fat mass variable and could be indirectly interpreted as total body weight, suggesting over fitting.

Similar to the Body Measurements Model, this model also considers `fat_percentage_category` to be the estimand. According to the study [], the two sexes have different thresholds for different ages, suggesting linkages to different metabolic needs based on different phases of life. Hence, `fat_percentage_category` is initially classified using `gender`, `age` and `fm`.

@@

Assumptions:

This model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.

@@

This model is also built on data that was generated by the ABS [] simulator so it may be inherently working with an error. While this is an indirect form of observed data, the original data set claims [x] accuracy. However, the error that it could cause in this paper is not accounted for and is considered the true measurement of the test subject.

\newpage
# Results
Our results are summarized in @tbl-measurementmodelresults and @tbl-bodymass-modelresults. Our model's results and interpretation:

@@

# Models Validation {#sec-model-details}
